---
title: "BOCC Visualization"
author: "Connor Gibbs"
date: "6/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r, include = FALSE}
library(purrr)
library(dplyr)
library(igraph)
library(ggplot2)
source('AnalyzeResults/utilities/read_write.R')
source('AnalyzeResults/utilities/wrangle.R')
source('AnalyzeResults/utilities/plot.R')
```

## Data

```{r}
bocc <- read_BOCCResults(path = 'Results/BOCC_Results')

stats2plot <-
  c(
    'cluster_size',
    'gene_ratio',
    'HPO_ratio',
    'num_sig_go_enrichment_terms',
    'max_norm_cell_type_specificity',
    'max_norm_cell_type_comma_sep_string',
    'max_norm_disease_specificity',
    'max_norm_disease_comma_sep_string'
  )

df <- bocc %>%
  create_stats_df(., stats2plot) %>%
  dplyr::mutate(., dplyr::across(c('gene_ratio', 'HPO_ratio'), ~round(.x, 3)))
```

## Cluster Sizes

```{r}
df %>%
  group_by(method) %>%
  summarize(
    .,
    n = length(cluster_id),
    min_m = min(cluster_size),
    mean_m = mean(cluster_size),
    sd_m = sd(cluster_size),
    max_m = max(cluster_size),
    sum_m = sum(cluster_size)
  ) 
  # %>%
  # kableExtra::kable(.,
  #                   format = 'latex',
  #                   digits = 2,
  #                   booktabs = TRUE) %>%
  # kableExtra::kable_styling(position = 'center') %>%
  # kableExtra::add_header_above(., list(' ' = 2, 'cluster size' = 5))
```

### Cell and Disease Specificity

On a natural scale...

```{r}
p1 <- ggplot(
  df %>%
    dplyr::filter(., cluster_size > 1),
  aes(
    x = max_norm_cell_type_specificity,
    y = max_norm_disease_specificity,
    color = method,
    label = cluster_id,
    label2 = cluster_size,
    label3 = num_sig_go_enrichment_terms,
    label4 = gene_ratio,
    label5 = HPO_ratio,
    label6 = max_norm_cell_type_comma_sep_string,
    label7 = max_norm_disease_comma_sep_string,
  )
) +
  geom_point(alpha = 0.5, position = 'jitter') +
  scale_color_brewer('method', palette = 'Set1') +
  labs(x = 'cell type specificity', y = 'disease type specificity') +
  theme_bw() +
  theme(legend.position = 'bottom',
        legend.direction = 'horizontal',
        legend.box = 'vertical')
p1
ggsave('AnalyzeResults/figs/bocc_results.png', width = 5, height = 5)
```

On a log-log scale...

```{r}
p2 <- p1 + 
  scale_x_log10(
    bquote(~log[10]('cell specificity')),
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    minor_breaks = log10_minor_break(),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  scale_y_log10(
    bquote(~log[10]('disease specificity')),
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    minor_breaks = log10_minor_break(),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  annotation_logticks()
p2
ggsave('AnalyzeResults/figs/bocc_results_loglog.png', width = 5, height = 5)
```

As a widget...

```{r}
p1 <- p1 +
  theme(legend.position = 'bottom')
plotly::ggplotly(p1)
```

### Deep Dive

We'll look into the communities with high specificity ($> 0.50$).

```{r}
deep_dive <- df %>%
  dplyr::filter(., (max_norm_cell_type_specificity > 0.5 | 
                  max_norm_disease_specificity > 0.5) & cluster_size > 1) %>%
  dplyr::select(., method:cluster_size, max_norm_cell_type_specificity, 
                max_norm_disease_specificity, dplyr::everything())
DT::datatable(deep_dive, options = list(pageLength = 5))

deep_dive %>%
  dplyr::select(., method, cluster_id) %>%
  readr::write_csv(., 'Results/promising_clusters_june_29_2021.csv')
```

#### Walktrap

I'll be investigating the cluster's identified by `cluster_id` `r sort(deep_dive$cluster_id[stringr::str_detect(deep_dive$method, 'walktrap')])`.

```{r, include=FALSE}
# data
comms <- 'Results/Communities' %>%
  read_Communities() %>%
  dplyr::bind_rows()

G <-
  read_graph(edgelist = 'Data/HPO_String_edgelist_june_22_2021.numbered.tsv',
             nodes = 'Data/HPO_String_edgelist_june_22_2021.nodenames.tsv',
             communities = 'Results/Communities')

# function to summarize community
summarize_cluster <- function(G, method, cluster_id, plot = FALSE) {
  mems <- igraph::V(G)$communities %>%
    dplyr::bind_rows() %>%
    dplyr::filter(., stringr::str_detect(method, !!method), cluster_id == !!cluster_id) %>%
    dplyr::pull(node)
  
  sG <- igraph::induced.subgraph(G, vids = V(G)[which(V(G)$node %in% mems)])
  
  if(plot){
    if(igraph::vcount(sG) < 1000){
      plot_graph(sG, legend = FALSE, vertex.label.color = 'black')
    }else{
      warning('No plot produced. Too many nodes.')
    }
  }
  
  return(list(sG = sG,
              n = vcount(sG),
              m = ecount(sG)))
}
```

```{r}
ids <- sort(deep_dive$cluster_id[stringr::str_detect(deep_dive$method, 'walktrap')])
results <- list()
results$walktrap <- map(ids, ~summarize_cluster(G = G, method = 'walktrap', cluster_id = .x, plot = TRUE))
```

#### Infomap

I'll be investigating the clusters identified by `cluster_id` `r sort(deep_dive$cluster_id[stringr::str_detect(deep_dive$method, 'infomap')])`.

```{r}
ids <- sort(deep_dive$cluster_id[stringr::str_detect(deep_dive$method, 'infomap')])
results$infomap <- map(ids, ~summarize_cluster(G = G, method = 'infomap', cluster_id = .x, plot = TRUE))
```