---
title: "Meeting for 08/03"
author: "Connor Gibbs"
date: "7/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(purrr)
library(dplyr)
library(igraph)
library(ggplot2)
library(igraphaid)
```

```{r, include=FALSE}
source('AnalyzeResults/utilities/read_write.R')
source('AnalyzeResults/utilities/eval.R')
source('AnalyzeResults/utilities/plot.R')
```

## Plot Communities Gained

I will plot the communities detected within the weighted graph that have a maximum Jaccard measure of 0. This means they do not correspond with any community identified with the unweighted graph.

I'll start by reading in the communities of interest. This file is created in `AnalyzeResults/rmd/meeting_20210727.Rmd`.

```{r, echo = FALSE}
gl <-
  readr::read_rds('Data/WeightedPhenotypicAbnormality/gain_loss_weighted2unweighted.rds')
vgl <- list()
vgl$gain <- purrr::map(gl$gain$members, ~.x$comm)
vgl$loss <- purrr::map(gl$loss$members, ~.x$comm)
rm(gl)
```

Now, I'll read in the graph.

```{r}
G <-
  read_graph(edgelist = 'Data/PhenotypicAbnormality/String_HPO_2015.phenotypic_branch.numbered.edgelist.txt',
             nodes = 'Data/PhenotypicAbnormality/String_HPO_2015.phenotypic_branch.nodenames.txt',
             communities = NULL)
igraph::is_weighted(G)
```

Now, I'll plot gained communities.

```{r}
for(i in 1:length(vgl$gain)){
  png(
    filename = stringr::str_c(
      'AnalyzeResults/figs/WeightedPhenotypicAbnormality/gain/',
      'gained-comm-',
      i,
      '.png'
    ),
    width = 5,
    height = 5,
    units = 'in',
    res = 100
  )
  plot_subgraph(G, vids = vgl$gain[[i]], order = 1, 
                vertex.size = 6, vertex.label.cex = 0.8, 
                layout = igraph::layout_with_fr)
  dev.off()
}
```

Now, I'll plot the lost communities.

```{r}
for(i in 1:length(vgl$loss)){
  png(
    filename = stringr::str_c(
      'AnalyzeResults/figs/WeightedPhenotypicAbnormality/loss/',
      'lost-comm-',
      i,
      '.png'
    ),
    width = 5,
    height = 5,
    units = 'in',
    res = 100
  )
  plot_subgraph(G, vids = vgl$loss[[i]], order = 1, 
                vertex.size = 6, vertex.label.cex = 0.8, 
                layout = igraph::layout_with_fr)
  dev.off()
}
```