---
title: "Meeting for 7/27"
author: "Connor Gibbs"
date: "7/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(purrr)
library(dplyr)
library(igraph)
library(ggplot2)
library(igraphaid)
```

```{r, include=FALSE}
source('AnalyzeResults/utilities/read_write.R')
source('AnalyzeResults/utilities/eval.R')
source('AnalyzeResults/utilities/plot.R')
```

## Compare Unweighted Walktrap to Weighted Walktrap

First, I will read in the weighted graph.

```{r}
G <-
  read_graph(edgelist = 'Data/WeightedPhenotypicAbnormality/String_HPO_2015.weighted.phenotypic_branch.numbered.edgelist.txt',
             nodes = 'Data/WeightedPhenotypicAbnormality/String_HPO_2015.weighted.phenotypic_branch.nodenames.txt',
             communities = NULL)
igraph::is_weighted(G)
```

Now, I will run the walktrap algorithm using the edge weights, keeping only communities with at least three members.

```{r, echo=FALSE}
# wt_comms <- igraph::cluster_walktrap(G)
wt_comms <- readr::read_rds('Results/communities/PhenotypicAbnormality/walktrap.String_HPO_2015.weighted.phenotypic_branch.coms')
wt_comms_list <- wt_comms$igraph_comms$membership %>%
  data.frame(node = wt_comms$igraph_comms$names, community = .) %>%
  dplyr::arrange(., community, node) %>%
  dplyr::group_split(., community) %>%
  purrr::map(., ~ .x$node) %>%
  purrr::keep(., ~ length(.x) >= 3)
# readr::write_rds(
#   list(comms = wt_comms_list, igraph_comms = wt_comms),
#   'Results/communities/PhenotypicAbnormality/walktrap.String_HPO_2015.weighted.phenotypic_branch.coms',
#   compress = 'gz'
# )
```

Now, I will pull in detected communities from the unweighted graph.

```{r, echo=FALSE}
## unweighted communities
unwt_comms_list <- 'Results/communities/PhenotypicAbnormality' %>%
  read_Communities() %>%
  .$walktrap.String_HPO_2015 %>%
  dplyr::group_split(., cluster_id) %>%
  purrr::map(., ~.x$name) %>%
  purrr::keep(., ~length(.x) >= 3)
```
Finally, I'll take a look at the number of communities and the distribution of community sizes:

```{r, echo=FALSE}
# number of weighted communities
length(wt_comms_list)
# number of unweighted communities
length(unwt_comms_list)

dt <- dplyr::bind_rows(
  data.frame(method = 'weighted', 
             nmembers = purrr::map_dbl(wt_comms_list, ~length(.x)), 
             id = 1:length(wt_comms_list)),
  data.frame(method = 'unweighted', 
             nmembers = purrr::map_dbl(unwt_comms_list, ~length(.x)),
             id = 1:length(unwt_comms_list))
) %>%
  dplyr::mutate(., label = ifelse(nmembers > 5000, id, ''))

ggplot(data = dt, aes(x = method, y = nmembers, color = method)) +
  geom_violin() +
  geom_jitter(alpha = 0.6, position = position_jitter(0.2)) +
  geom_text(label = dt$label) +
  scale_y_log10(
    bquote(~log[10]('number of members')),
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    minor_breaks = log10_minor_break(),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  annotation_logticks() +
  scale_color_brewer('', palette = 'Set1') +
  labs(title = 'Number of Members in Walktrap Communities') +
  theme_bw() +
  theme(legend.position = 'none')
ggsave('AnalyzeResults/figs/weighted-number-of-members.png', height = 4, width = 6, units = 'in')
```

And compare the overlap in the identified communities...

```{r}
jaccard <- avg_jaccard(A = wt_comms_list, B = unwt_comms_list)
jaccard$avg # average over all pairs and matching direction

# for each weighted community, find the unweighted community that maximized jaccard
dt_wtu <- jaccard$grid %>%
  dplyr::group_by(., A) %>%
  dplyr::slice_max(., jaccard, with_ties = FALSE)

dt_wtu %>%
  DT::datatable(
  .,
  extensions = 'Buttons',
  options = list(
    dom = 'Blfrtip',
    buttons = c('copy', 'csv'),
    lengthMenu = list(c(10, 25, 50, -1),
                      c(10, 25, 50, "All"))
  )
) %>%
  DT::formatRound(., columns = 'jaccard', digits = 2)

p1 <- ggplot2::ggplot(data = dt_wtu, aes(x = jaccard)) +
  geom_density() +
  geom_rug(data = dt_wtu, aes(x = jaccard, y = 0), sides = 'b', position = 'jitter', alpha = 0.7) +
  labs(x = 'maximum jaccard', y = 'density', 
       title = 'Weighted to Unweighted') +
  lims(y = c(0, 1.2)) +
  theme_bw() +
  theme(legend.position = 'none')

# for each unweighted community, find the weighted community that maximized jaccard
dt_utw <- jaccard$grid %>%
  dplyr::group_by(., B) %>%
  dplyr::slice_max(., jaccard, with_ties = FALSE)

dt_utw %>%
  DT::datatable(
  .,
  extensions = 'Buttons',
  options = list(
    dom = 'Blfrtip',
    buttons = c('copy', 'csv'),
    lengthMenu = list(c(10, 25, 50, -1),
                      c(10, 25, 50, "All"))
  )
) %>%
  DT::formatRound(., columns = 'jaccard', digits = 2)

p2 <- ggplot2::ggplot(data = dt_utw, aes(x = jaccard)) +
  geom_density() +
  geom_rug(data = dt_utw, aes(x = jaccard, y = 0), sides = 'b', position = 'jitter', alpha = 0.7) +
  labs(x = 'maximum jaccard', y = 'density', 
       title = 'Unweighted to Weighted') +
  lims(y = c(0, 1.2)) +
  theme_bw() +
  theme(legend.position = 'none')

ggpubr::ggarrange(p1, p2) %>%
  ggpubr::annotate_figure(.,
                          top = ggpubr::text_grob(
                            'Distribution of Maximal Jaccard',
                            family = "sans",
                            size = 14,
                            face = 'bold'
                          ))
ggsave('AnalyzeResults/figs/distn-of-jaccard.png', width = 6, height = 4, units = 'in')
rm(p1, p2)
```

This leads me to the following question: which communities are identified via the weighted graph but not the unweighted graph and at what cost?

```{r}
n <- nrow(dt_wtu)
dt_wtu %>%
  dplyr::filter(., jaccard %in% c(0, 1)) %>%
  dplyr::group_by(., jaccard) %>%
  dplyr::summarise(., n = n(), freq = n()/!!n)

dt <- dt_wtu %>%
  dplyr::filter(., jaccard == 0) %>%
  dplyr::pull(A) %>%
  purrr::map_dfr(., ~ {
    comm <- as.vector(unlist(wt_comms_list[.x]))
    tibble::tibble(id = .x,
                   comm = comm,
                   n = length(comm),
                   nhpo = sum(stringr::str_sub(comm, 1, 3) == 'HP:'),
                   ngene = sum(stringr::str_sub(comm, 1, 3) != 'HP:')) %>%
      dplyr::mutate(., perc_gene = ngene/n, 
                    perc_hpo = nhpo/n) %>%
      tidyr::nest(., members = comm)
  })

# what is gained by using a weighted graph?
gained <- dplyr::filter(dt, perc_gene > 0, perc_hpo > 0)
DT::datatable(gained)

dt <- 1:nrow(dt_utw) %>%
  setdiff(., dt_wtu$B) %>%
  purrr::map_dfr(., ~ {
    comm <- as.vector(unlist(unwt_comms_list[.x]))
    tibble::tibble(id = .x,
                   comm = comm,
                   n = length(comm),
                   nhpo = sum(stringr::str_sub(comm, 1, 3) == 'HP:'),
                   ngene = sum(stringr::str_sub(comm, 1, 3) != 'HP:')) %>%
      dplyr::mutate(., perc_gene = ngene/n, 
                    perc_hpo = nhpo/n) %>%
      tidyr::nest(., members = comm)
  })

# what is lost by using a weighted graph?
lost <- dplyr::filter(dt, perc_gene > 0, perc_hpo > 0)
DT::datatable(lost)
readr::write_rds(list(gain = gained, loss = lost), 'Data/WeightedPhenotypicAbnormality/gain_loss_weighted2unweighted.rds')
```

Now, which communities are identified via the unweighted graph but not the weighted graph and at what cost?

```{r}
n <- nrow(dt_utw)
dt_utw %>%
  dplyr::filter(., jaccard %in% c(0, 1)) %>%
  dplyr::group_by(., jaccard) %>%
  dplyr::summarise(., n = n(), freq = n()/!!n)

# what is gained by not using a weighted graph?
## Nothing... Could look at jaccard near 0
```
