---
title: "Meeting for 6/22"
author: "Connor Gibbs"
date: "6/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(purrr)
library(dplyr)
library(igraph)
```

```{r, include=FALSE}
source('AnalyzeResults/utilities/read_write.R')
```

## Identified Communities

The result is a list of size $m$ where $m$ is the number of methods used. Each element of the list maintains three columns: `method`, `community`, and `member`. The method describes the clustering method. The community is the identifier created in Python, and the member describes the node, be it a gene or HPO. The elements of the list can be joined together with `dplyr::bind_rows()`. If a member has multiple rows, that means it is a member of multiple communities for at least one overlapping community detection method.

```{r}
comms <- read_Communities('Results/Communities')
head(comms$cesna)
```

## BOCC Results

The result is a list of size $m$ where $m$ is the number of methods used. Each element of the list is a data frame where comma separated elements are saved as a list. Each row summarizes a particular community with an identifier given by `cluster_id`. 

```{r}
bocc <- read_BOCCResults('Results/BOCC_Results')
head(bocc$cesna)
```

## Graph

The result is an `igraph` object where there are node attributes for the name of the node (`name`), the Python node ID (`node`), indicator of whether the node is a gene (`gene`), and a data frame of the corresponding communities (`communities`).

```{r}
G <-
  read_graph(edgelist = 'Data/HPO_String_edgelist_june_22_2021.numbered.tsv',
             nodes = 'Data/HPO_String_edgelist_june_22_2021.nodenames.tsv',
             communities = 'Results/Communities')
vertex_attr_names(G)

V(G)[V(G)$node %in% c(7, 12000)]$name
V(G)[V(G)$node %in% c(7, 12000)]$gene
V(G)[V(G)$node %in% c(7, 12000)]$communities
```

### Discussion

Why is the graph a multi-graph? Do the multiple edges tell us anything biologically meaningful or just an artifact of how the graph is constructed?

*solved*

```{r}
el <- 'Data/HPO_String_edgelist_june_22_2021.numbered.tsv' %>%
  readr::read_delim(., delim = '\t', col_names = FALSE) %>%
  purrr::set_names(., c('from', 'to'))

# sort the edges and remove duplicates
el_sort <- swap_if(el, c('from', 'to'), c('from', 'to'))
nrow(dplyr::distinct(el_sort))
nrow(el)
```

Why are there self loops? Are these loops biologically meaningful or an artifact of the data preparation process?

```{r}
dplyr::filter(el, from == to)
```